Estimation of the road traffic sound levels based on Non-Negative Matrix Factorization

### 2017-12-19 article/articleApplied.tex
following update

%or by generating a soft mask, $\mathbf{M_s}$,
%
%\begin{equation} \label{eq:definitionMask}
%\mathbf{M_s} = \frac{\left[\mathbf{W} \mathbf{H}\right]_{traffic}}{\mathbf{W H}},
%\end{equation}
%
%which allows to determine the spectrogram of the traffic component, $ \mathbf{\tilde{V}}_{traffic}$
%
%\begin{equation}\label{eq:separation}
%\mathbf{\tilde{V}}_{traffic} = \mathbf{M_s}.\mathbf{V}.
%\end{equation}

@@

Supervised and semi-supervised NMF are performed for 400 iterations which is sufficiently enough to get a stabilized reconstruction. ThC-NMF is performed on a lower number of iteration (60) to prevent $\mathbf{W}$ to not deviate to much from the initial dictionary. The spectrogram $\mathbf{V}$ and the dictionary $\mathbf{W}$ are expressed on two different formats: with a linear frequency scale ($\Delta f \approx 10.8$ Hz) and with third octave bands (29 bands). Theses two methods are considered to compare a fine grain approach (the linear scale) with a coarser one (the third octave bands) as it reduces the number of frequency bins and allows to reduces the number of bands in the high frequencies where the traffic component is less present.  Furthermore, in the case of the linear frequency scale and for supervised and semi-supervised NMF, $\mathbf{V}$ and $\mathbf{W}$ are filtered at the frequencies $f_c$ in order to focused the reconstruction of the signal of the low frequency bins. Nevertheless, if NMF is performed with filtered elements ($\mathbf{V}_{f_c}$ and $\mathbf{W}_{f_c}$) to determine $\mathbf{H}_{f_c}$, the traffic signal reconstruction is done with the original dictionary $\mathbf{W}$, as

@@

With the third octave frequency scale, only the case $f_c = 20$ kHz is applied. Finally, for the ThC NMF, the threshold is define between 0.20 and 0.70 with a 0.01 step. Table \ref{tab:estimation_factors} summarizes the parameters and the related values.

@@

\begin{table*}[]
\centering
\begin{tabular}{@{}cccccccc@{}}
\toprule
$\mathbf{K}$ & $\mathbf{w_t}$ (s) & $\mathbf{f_c}$ (kHz) &\begin{tabular}[c]{@{}c@{}}\textbf{spectral} \\ \textbf{representation}\end{tabular} &  $\mathbf{\beta}$ & \textbf{method} & $\mathbf{t}$ & \textbf{$MAE$ (dB)} \\ \midrule
 &  & 20 &  &  & filter &  & 4.69 ($\pm$ 4.52) \\ 
 &  & 0.5 &  &  & filter &  & 2.89 ($\pm$ 2.84) \\ \hline \hline
25 & all & 0.5 & spectra & 1 & supervised NMF &  & \textbf{2.76 ($\pm$ 2.86)} \\ 
50 & 0.5 & 0.5 & spectra & 2 & supervised NMF &  & \textbf{2.56 ($\pm$ 2.48)} \\ 
50 & 0.5 &  20 & third octave & 1 & supervised NMF &  & 3.44 ($\pm$ 3.70) \\ 
50 & 0.5 & 20 & third octave & 2 & supervised NMF &  & 3.02 ($\pm$ 3.33) \\ \hline \hline
100 & 0.5 &  20 & spectra & 1 & semi-supervised NMF &  & \textbf{2.36 ($\pm$ 1.23)} \\ 
100 & 0.5 & 20 & spectra & 2 & semi-supervised NMF &  & \textbf{2.34 ($\pm$ 1.35)} \\ 
100 & 0.5  & 20 & third octave & 1 & semi-supervised NMF &  & \textbf{2.38 ($\pm$ 1.26)} \\ 
100 & 0.5 & 20 & third octave & 2 & semi-supervised NMF &  & \textbf{2.43 ($\pm$ 1.43)} \\ \hline \hline
 &  &  & spectra & 1 & ThC NMF &  &  \\ 
100 & all & 20 & spectra & 2 & ThC NMF & 0.37 & \textbf{2.22 ($\pm$ 2.37)} \\ 
100 & all & 20 & third octave & 1 & ThC NMF & 0.57 & \textbf{2.19 ($\pm$ 2.01)} \\ 
100 & all & 20 & third octave & 2 & ThC NMF & 0.54 & \textbf{2.16 ($\pm$ 2.24)}\\ \bottomrule
\end{tabular}
\caption{Best results according to $\beta$, method and  spectral representation}
\label{tab:results}
\end{table*}

@@

The results of different versions of NMF are summarized next to the filter results with, in bold, the NMF errors lower than the baseline. With the exception of supervised NMF with third octave spectral representation, all NMF have a lower mean error than the baseline. However, the best combination is got with a ThC NMF with a third octave spectral representation, $\beta$ = 2 and a threshold $t$ = 0.54. For each method (supervised, semi-supervised and ThC NMF), the best scenario is detailed according the sub-classes and the $TIR$ (Figure \ref{fig:TIR_class_filter}, \ref{fig:TIR_class_sup}, \ref{fig:TIR_class_semi} and \ref{fig:W_ThC_NMF}).

@@

\begin{figure*}[t!]
    \centering
    \begin{subfigure}[t]{0.45\textwidth}
        \centering
        \includegraphics[width=\linewidth]{../image/filter_bar.pdf}
        \caption{filter method with $f_c$ = 500 Hz}
        \label{fig:TIR_class_filter}
    \end{subfigure}%
    \hfill
    \begin{subfigure}[t]{0.45\textwidth}
        \centering
        \includegraphics[width=\linewidth]{../image/sup_bar.pdf}
        \caption{Supervised NMF, $\beta$ = 2, linear spectral representation, $f_c$ = 0.5 kHz}
                \label{fig:TIR_class_sup}
    \end{subfigure}

@@

The gain of the semi-supervised approach is most of all significant for these low $TIR$ where the add of a mobile part, $\mathbf{W_r}$, into $\mathbf{W}$ allows to take into account the other predominant sound sources (example in Figure \ref{fig:Y_alert-12}) while supervised NMF is constrained to use traffic elements to reduce the distance/divergence between $\mathbf{V}$ and $\mathbf{WH}$. This improvement, according to supervised NMF results, can be seen in Figure \ref{fig:TIR_class_semi} where the error on each sub class for $TIR = -12$ and $TIR$ = -6 are much lower.  Nevertheless, these degrees of freedom are restrictive for high $TIR$. Indeed, without constraint semi-supervised NMF is free to include traffic components in $W_r$ decreasing the rebuilt of the traffic component (Figure \ref{fig:Y_alert_12}). The error are then increasing for theses $TIR$. In the opposite, supervised NMF has to use the \textit{traffic} element present in $W$ which improve the performances.

\begin{figure}[t!]
    \centering
    \begin{subfigure}[t]{0.45\textwidth}
        \centering
        \includegraphics[width=\linewidth]{../image/Y_alert_TPR_-12_01.pdf}
        \caption{$\mathbf{W_r}$ for an \textit{alert} scene at $TIR$ = -12}
        \label{fig:Y_alert-12}
    \end{subfigure}%
    \hfill
    \begin{subfigure}[t]{0.45\textwidth}
        \centering
        \includegraphics[width=\linewidth]{../image/Y_alert_TPR_12_01.pdf}
        \caption{$\mathbf{W_r}$ for an \textit{alert} scene at $TIR$ = 12}
		\label{fig:Y_alert_12}
    \end{subfigure}
\end{figure}

Finally, ThC NMF with a threshold fixed at $t = 0.54$ offers the most lower results. The dictionary update allows to the method to adapt to the different $TIR$ and to only keep the closest elements of the \textit{traffic} component \ref{fig:dist_-12_12}.

\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{../image/dist_-12_12.pdf}    
    \caption{Example of the similarity $\cos \theta$ for an sound mixture of the \textit{alert} sub-class for two $TIR$ and with the threshold $t$: $TIR = -12$ (a) and $TIR = 12$ (b)}
    \label{fig:dist_-12_12}
\end{figure}

The gain is significant on all $TIR$ with the exception of $TIR = 0$. On figure \ref{fig:ThC_error}, we can see the improvement of .
While supervised NMF have high errors on low $TIR$ for the sub-classes \textit{human} and \textit{transport}, ThC NMF succeed to have error inferior to 4 dB. The error for \textit{climate} and \textit{mechanics} are, on the other hand, similar to those of supervised NMF.

>>>>

@@

Supervised and semi-supervised NMF are performed for 400 iterations which is sufficiently enough to get a stabilized reconstruction. ThC-NMF is performed on a lower number of iteration (60) to prevent $\mathbf{W}$ to not deviate to much from the initial dictionary. The spectrogram $\mathbf{V}$ and the dictionary $\mathbf{W}$ are expressed on two different spectral representation: with a linear frequency scale ($\Delta f \approx 10.8$ Hz) and with third octave bands (29 bands). Theses two methods are considered to compare a fine grain approach (the linear scale) with a coarser one (the third octave bands) as it reduces the number of frequency bins and allows to reduces the number of bands in the high frequencies where the traffic component is less present.  Furthermore, in the case of the linear frequency scale and for supervised and semi-supervised NMF, $\mathbf{V}$ and $\mathbf{W}$ are filtered at the frequencies $f_c$ in order to focused the reconstruction of the signal of the low frequency bins. Nevertheless, if NMF is performed with filtered elements ($\mathbf{V}_{f_c}$ and $\mathbf{W}_{f_c}$) to determine $\mathbf{H}_{f_c}$, the traffic signal reconstruction is done with the original dictionary $\mathbf{W}$, as

@@

With the third octave spectral representation, only the case $f_c = 20$ kHz is applied. Finally, for the ThC NMF, the threshold is define between 0.20 and 0.70 with a 0.01 step. Table \ref{tab:estimation_factors} summarizes the parameters and the related values.

@@

\begin{table*}[]
\centering
\begin{tabular}{@{}cccccccc@{}}
\toprule
$\mathbf{K}$ & $\mathbf{w_t}$ (s) & $\mathbf{f_c}$ (kHz) &\begin{tabular}[c]{@{}c@{}}\textbf{spectral} \\ \textbf{representation}\end{tabular} &  $\mathbf{\beta}$ & \textbf{method} & $\mathbf{t}$ & \textbf{$MAE$ (dB)} \\ \midrule
 &  & 20 &  &  & filter &  & 4.69 ($\pm$ 4.52) \\ 
 &  & 0.5 &  &  & filter &  & 2.89 ($\pm$ 2.84) \\ \hline \hline
25 & all & 0.5 & spectra & 1 & supervised NMF &  & \textbf{2.76 ($\pm$ 2.86)} \\ 
50 & 0.5 & 0.5 & spectra & 2 & supervised NMF &  & \textbf{2.56 ($\pm$ 2.48)} \\ 
50 & 0.5 &  20 & third octave & 1 & supervised NMF &  & 3.44 ($\pm$ 3.70) \\ 
50 & 0.5 & 20 & third octave & 2 & supervised NMF &  & 3.02 ($\pm$ 3.33) \\ \hline \hline
100 & 0.5 &  20 & spectra & 1 & semi-supervised NMF &  & \textbf{2.36 ($\pm$ 1.23)} \\ 
100 & 0.5 & 20 & spectra & 2 & semi-supervised NMF &  & \textbf{2.34 ($\pm$ 1.35)} \\ 
100 & 0.5  & 20 & third octave & 1 & semi-supervised NMF &  & \textbf{2.38 ($\pm$ 1.26)} \\ 
100 & 0.5 & 20 & third octave & 2 & semi-supervised NMF &  & \textbf{2.43 ($\pm$ 1.43)} \\ \hline \hline
50 & all & 5 & spectra & 1 & ThC NMF & 0.39 & \textbf{2.23 ($\pm$ 2.21)} \\ 
50 & all & 2 & spectra & 2 & ThC NMF & 0.37 & \textbf{2.17 ($\pm$ 2.28)} \\ 
100 & all & 20 & third octave & 1 & ThC NMF & 0.57 & \textbf{2.19 ($\pm$ 2.01)} \\ 
100 & all & 20 & third octave & 2 & ThC NMF & 0.54 & \textbf{2.16 ($\pm$ 2.24)}\\ \bottomrule
\end{tabular}
\caption{Best results according to $\beta$, method and  spectral representation}
\label{tab:results}
\end{table*}

@@

The results of different versions of NMF are summarized next to the filter results with the NMF errors lower than the baseline in bold. With the exception of supervised NMF with third octave spectral representation, all NMF have a lower mean error than the baseline. The best combination is got with a ThC NMF with a third octave spectral representation, $\beta$ = 2 and a threshold $t$ = 0.54.

Between the dictionary parameters ($K$ and $w_t$) and the cut-off frequency $f_c$, no parameters is common to all the settings. For each method (supervised, semi-supervised and ThC NMF), the best scenario is detailed according the sub-classes and the $TIR$ (Figure \ref{fig:TIR_class_filter}, \ref{fig:TIR_class_sup}, \ref{fig:TIR_class_semi} and \ref{fig:W_ThC_NMF}).

@@

\begin{figure*}[t]
    \centering
    \begin{subfigure}[t]{0.45\textwidth}
        \centering
        \includegraphics[width=\linewidth]{../image/filter_bar.pdf}
        \caption{Frequency low-pass filter with $f_c$ = 500 Hz}
        \label{fig:TIR_class_filter}
    \end{subfigure}%
    \hfill
    \begin{subfigure}[t]{0.45\textwidth}
        \centering
        \includegraphics[width=\linewidth]{../image/sup_bar.pdf}
        \caption{Supervised NMF, $\beta$ = 2, linear spectral representation, $f_c$ = 0.5 kHz}
                \label{fig:TIR_class_sup}
    \end{subfigure}

@@

The gain of the semi-supervised approach is most of all significant for these low $TIR$ where the add of a mobile part, $\mathbf{W_r}$, into $\mathbf{W}$ allows to take into account the other predominant sound sources (example in Figure \ref{fig:Y_alert-12}) while supervised NMF is constrained to use traffic elements to reduce the distance/divergence between $\mathbf{V}$ and $\mathbf{WH}$. This improvement, according to supervised NMF results, can be seen in Figure \ref{fig:TIR_class_semi} where the error on each sub class for $TIR = -12$ and $TIR$ = -6 are much lower.  Nevertheless, these degrees of freedom are restrictive for high $TIR$. Indeed, without constraint semi-supervised NMF is free to include traffic components in $\mathbf{W_r}$ decreasing the rebuilt of the traffic component (Figure \ref{fig:Y_alert_12}). The error are then increasing for theses $TIR$. In the opposite, supervised NMF has to use the \textit{traffic} element present in $W$ which improve the performances. \\

\begin{figure*}[t]
    \centering
    \begin{subfigure}[t]{0.45\textwidth}
        \centering
        \includegraphics[width=\linewidth]{../image/Y_alert_TPR_-12_01.pdf}
        \caption{$\mathbf{W_r}$ for an \textit{alert} scene at $TIR$ = -12}
        \label{fig:Y_alert-12}
    \end{subfigure}%
    \hfill
    \begin{subfigure}[t]{0.45\textwidth}
        \centering
        \includegraphics[width=\linewidth]{../image/Y_alert_TPR_12_01.pdf}
        \caption{$\mathbf{W_r}$ for an \textit{alert} scene at $TIR$ = 12}
		\label{fig:Y_alert_12}
    \end{subfigure}
\end{figure*}

Finally, ThC NMF with a hard threshold fixed at $t = 0.54$ offers the lowest results. The dictionary update allows to the method to adapt to the different $TIR$ and to only keep the closest elements of the \textit{traffic} component \ref{fig:dist_-12_12}.

\begin{figure}[t]
    \centering
    \includegraphics[width=\linewidth]{../image/dist_-12_12.pdf}    
    \caption{Example of the similarity $\cos \theta$ for an sound mixture of the \textit{alert} sub-class for two $TIR$ and with the threshold $t$: $TIR = -12$ (a) and $TIR = 12$ (b)}
    \label{fig:dist_-12_12}
\end{figure}

The gain is significant on all $TIR$ with the exception of $TIR = 0$. On Figure \ref{fig:TIR_class_thc}, we can see the improvement bringing by this approach. For low $TIR$, only a reduce number of $\mathbf{W}$ are considered as traffic components, the performance on \textit{alert}, \textit{animals}, \textit{humans} and \textit{tranportation} are then similar to semi-supervised NMF. However, for \textit{climate} and \textit{mechanics}, the error are important as in supervised NMF. On high $TIR$, the similarity of the initial dictionary and $\mathbf{W}$ allows to keep more elements. In these cases, ThC NMF is the most powerful of the proposed methods. 
Consequently, if the average error is lower than semi-supervised NMF, with good performances on high $TIR$, it is a more variable method.
